---
title: "LAB 1"
author: "2048015 - MANOJ KUMAR"
date: "22/01/2021"
output: word_document
---

### 1. Define the object "myobject" and assign the vector 1:10 in at least 3 different ways.

```{r}
# To create a vector with numerical values in a sequence, use the : operator and seq() function can be used. 

# Type - 1
myobject <- 1:10
myobject
```
```{r}
# Type - 2
myobject <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
myobject
```
```{r}
# Type - 3
myobject <- seq(from = 1,
                to   = 10,
                by   = 1)
myobject
```


### 2. Get the sum of your object.

```{r}
# To get the sum of the object sum() function can be used.

sum_myobject <- sum(myobject)
sum_myobject
```

### 3. Create the following vector by using the paste function.

    [1] "R is great 4 and I will love it"
    [2] "R is great 7 and I will love it"
    [3] "R is great 45 and I will love it"
   
```{r}
n = paste0(c(4,7,45))
paste("R is great",n,"and I will love it.")
```
   
### 4. Vector of 1,2,3, repeat the vector to get 11 x 1, 10 x 2, and 10 x3 

```{r}
Vector = c(1,2,3)
Vector_1 = rep(c(11,10,10),
               times = Vector)
Vector_1
```

### 5. What is the value of this vector on position 7?

```{r}
Vector_1[7]
```

### 6. Repeat the string “Hello R” thrice.

```{r}
rep("Hello R", 3)
```

### 7. Repeat the first  element of a vector twice and the second element of the vector thrice.

  Eg: 1  1  2 2 2
 
```{r}
# Repeating each independently

# case 1
vector <- rep(c(5, 7), times = c(2, 3))
vector
```
```{r}
# case 2
vector = c(3,5,7)
d = length(vector)
l = 2:(d+1)

rep(vector,times = l)
```

           
### 8. Create a matrix of two rows and three columns.

```{r}
matrix_a <- matrix(1:6,  # Taking sequence of elements using : operator 
                   2,    # No of rows 
                   3     # No of columns 
                   )     # By default matrices are in column-wise order
             
matrix_a
```

### 9. Create a matrix 3X3 by row-wise.

```{r}
matrix_b <- matrix(1:9,         
                   nrow  = 3,   
                   ncol  = 3,     
                   byrow = TRUE 
                   # This parameter decides how to arrange the matrix 
                   )
matrix_b
```

### 10. Build a vectors of random values with the sample() function.

    Find the min(),max(),range(),length(),sum(),prod(),mean(),var().
  
```{r}
sample_vector <- sample(15)
sample_vector
```
```{r}
min(sample_vector)
max(sample_vector)
range(sample_vector)
length(sample_vector)
sum(sample_vector)
prod(sample_vector)
mean(sample_vector)
var(sample_vector)
```

