---
title: "Regression Analysis"
author: "MANOJ KUMAR - 2048015"
date: "20/02/2021"
output: word_document
---


## 1.Install the package “titanic”.
```{r}
# install.packages("titanic")
```

## 2.Load Titanic library to get the dataset
```{r}
# Load Titanic library
library(titanic)
```

```{r}
# Load the dataset
data("titanic_train")
data("titanic_test")
```

## 3.Set Survived column for test data to NA.
    #Note: titanic_test$Survived <- NA
```{r}
## Setting Survived column for test data to NA
titanic_test$Survived <- NA
```

## 4.Combine the  Training and Testing dataset.
    #Note: complete_data <- rbind(titanic_train, titanic_test)
```{r}
complete_data <- rbind(titanic_train, titanic_test)
```

## 5.Get the data structure.
```{r}
# Check data structure
str(complete_data)
```

## 6. Check for any missing values in the data.
```{r}
# Total missing count
colSums(is.na(complete_data))
```

## 7.Check for any empty values.
```{r}
colSums(complete_data=='')
```

```{r}
# Checking for Empty values
is.null(complete_data)
```

## 8.Check number of unique values for each column to find out which column we can convert to factors.
```{r}
apply(complete_data,2,function(x) length(unique(x)))
```

```{r}
# sapply() function takes list, vector or data frame as input and gives output in vector or matrix. 
# It is useful for operations on list objects and returns a list object of same length of original set.

sapply(complete_data, function(x) length(unique(x)))

```

## 9.Remove Cabin as it has very high missing values, passengerId, Ticket and Name are not required.
```{r}
# To remove a column from an R data frame

refined_data <- subset (complete_data, select = -c(Cabin, PassengerId, Ticket, Name))
head(refined_data)
```

## 10.Convert "Survived","Pclass","Sex","Embarked" to factors
```{r}
refined_data$Survived<-as.factor(refined_data$Survived)
refined_data$Pclass<-as.factor(refined_data$Pclass)
refined_data$Sex<-as.factor(refined_data$Sex)
refined_data$Embarked<-as.factor(refined_data$Embarked)
```

```{r}
str(refined_data)
```
```{r}
summary(refined_data)
```


## 11.Splitting training and test data.
```{r}
library("dplyr")
```

```{r}
df<-refined_data%>%
  filter(!is.na(Survived))
summary(df)
```

```{r}
library("mice")
```

```{r}
md.pattern(df)
```

```{r}
imputed_data<-mice(df,method = 'pmm',seed=50)
```
```{r}
summary(imputed_data)
```
```{r}
imputed_final<-complete(imputed_data)
summary(imputed_final)
```
```{r}
set.seed(42)

train_pts<-sample(1:nrow(imputed_final),0.75*nrow(imputed_final))

train_dataset <-imputed_final[train_pts,]
test_dataset <-imputed_final[-train_pts,]
```

```{r}
summary(train_dataset)
```

```{r}
Xtrain = subset(train_dataset,select=-c(Survived))
Ytrain = train_dataset$Survived
```
```{r}
library(caret)
```
```{r}
train_final <- upSample (subset(train_dataset,
                                select=-c(Survived)),
                         train_dataset$Survived)
summary(train_final)
```

## 12.Create a model.
```{r}
# The basic syntax for glm() function in logistic regression is −
#           glm(formula, data,family)

# formula is the symbol presenting the relationship between the variables.
# data is the data set giving the values of these variables.
# family is R object to specify the details of the model. It's value is binomial for logistic regression.

LogisticModel <- glm(Class ~., train_final, family = binomial(link='logit'))
LogisticModel
```

## 13.Visualize the model summary.
```{r}
# Model Summary
summary(LogisticModel)
```

## 14.Analyse the test of deviance using anova()
    #Note: anova(model, test="Chisq")
```{r}
# Using anova() to analyze the table of devaiance

anova(LogisticModel, test="Chisq")
```

## 15.Compute confusion matrix and ROC curve.
```{r}
## Predicting Test Data

data <- predict( LogisticModel, newdata = test_dataset, type="response")
pred_num <- ifelse(data > 0.5, 1, 0)

pred_data <- factor(pred_num, levels = c(0, 1))
actual_data <- test_dataset$Survived

confusionMatrix( data = actual_data, reference = actual_data)
```

```{r}
library(ROCR)
```

```{r}
prediction_obj <- prediction(as.numeric(pred_data), as.numeric(actual_data))
final_set <- performance(prediction_obj, "tpr", "fpr")
plot(final_set, colorize=TRUE)
```

