---
title: "Regression Analysis"
author: "MANOJ KUMAR - 2048015"
date: "20/02/2021"
output: word_document
---


## 1.Install the package “titanic”.
```{r}
# install.packages("titanic")
```

## 2.Load Titanic library to get the dataset
```{r}
# Load Titanic library
library(titanic)
```


```{r}
# Load the dataset
data("titanic_train")
data("titanic_test")
```

## 3.Set Survived column for test data to NA.
    #Note: titanic_test$Survived <- NA
```{r}
## Setting Survived column for test data to NA
titanic_test$Survived <- NA
```

## 4.Combine the  Training and Testing dataset.
    #Note: complete_data <- rbind(titanic_train, titanic_test)
```{r}
complete_data <- rbind(titanic_train, titanic_test)
```

## 5.Get the data structure.
```{r}
# Check data structure
str(complete_data)
```

## 6. Check for any missing values in the data.
```{r}
is.na(complete_data)
```

```{r}
# Total missing count
colSums(is.na(complete_data))
```

## 7.Check for any empty values.
```{r}
# Checking for Empty values

colSums(complete_data=='')
```

## 8.Check number of unique values for each column to find out which column we can convert to factors.
```{r}
sapply(complete_data, function(x) length(unique(x)))
```

```{r}
## Missing values imputation
complete_data$Embarked[complete_data$Embarked==""] <- "S"
complete_data$Age[is.na(complete_data$Age)] <- median(complete_data$Age,na.rm=T)
```

## 9.Remove Cabin as it has very high missing values, passengerId, Ticket and Name are not required.

```{r}
library(base)
```


```{r}
library(stats)
```


```{r}
library(dplyr)
```


```{r}
titanic_data <- complete_data %>% select(-c(Cabin, PassengerId, Ticket, Name))
```

## 10.Convert "Survived","Pclass","Sex","Embarked" to factors
```{r}
for (i in c("Survived","Pclass","Sex","Embarked")){
  titanic_data[,i]=as.factor(titanic_data[,i])
}
```

```{r}
# Create dummy variables for categorical variables
library(dummies)
```

```{r}
titanic_data <- dummy.data.frame(titanic_data, names=c("Pclass","Sex","Embarked"), sep="_")
```


## 11.Splitting training and test data.
```{r}
train <- titanic_data[1:667,]
test <- titanic_data[668:889,]
```

## 12.Create a model.
```{r}
## Model Creation
model <- glm(Survived ~.,family=binomial(link='logit'),data=train)
```

## 13.Visualize the model summary.
```{r}
# Model Summary
summary(LogisticModel)
```

## 14.Analyse the test of deviance using anova()
    #Note: anova(model, test="Chisq")
```{r}
## Using anova() to analyze the table of devaiance
anova(model, test="Chisq")
```

```{r}
## Predicting Test Data
result <- predict(model,newdata=test,type='response')
result <- ifelse(result > 0.5,1,0)
```

## 15.Compute confusion matrix and ROC curve.
```{r}
## Confusion matrix and statistics
library(ggplot2) 
```
```{r}
library(lattice)
```
```{r}
library(caret)
```

```{r}
confusionMatrix(data=result, reference=test$Survived)
```

```{r}
## ROC Curve and calculating the area under the curve(AUC)
library(ROCR)
```


```{r}
predictions <- predict(model, newdata=test, type="response")
ROCRpred <- prediction(predictions, test$Survived)
ROCRperf <- performance(ROCRpred, measure = "tpr", x.measure = "fpr")

plot(ROCRperf, colorize = TRUE, text.adj = c(-0.2,1.7), print.cutoffs.at = seq(0,1,0.1))
```
```{r}
auc <- performance(ROCRpred, measure = "auc")
auc <- auc@y.values[[1]]
auc
```

