---
title: "LAB7 REGRESSION TREE"
author: "SOUNDARYA G - 2048057"
date: "12/03/2021"
output:
  html_document:
    df_print: paged
---

### QUESTION 1.Download the dataset BOSTON.csv

```{r}
# Loading BOSTON dataset.

dataset <- read.csv("boston.csv")
head(dataset)
```
    
    - LON and LAT are the longitude and latitude of the center of the census tract.
    - MEDV is the median value of owner-occupied homes, measured in thousands of dollars.
    - CRIM is the per capita crime rate.
    - ZN is related to how much of the land is zoned for large residential properties.
    - INDUS is the proportion of the area used for industry.
    - CHAS is 1 if a census tract is next to the Charles River else 0 
    - NOX is the concentration of nitrous oxides in the air, a measure of air pollution.
    - RM is the average number of rooms per dwelling.
    - AGE is the proportion of owner-occupied units built before 1940.
    - DIS is a measure of how far the tract is from centres of employment in Boston.
    - RAD is a measure of closeness to important highways.
    - TAX is the property tax per $10,000 of value.
    - PTRATIO is the pupil to teacher ratio by town

```{r}
# structure of dataset
str(dataset)
```

**OBSERVARTION:**
     1.Character- TOWN
     2.Integer- TRACT,CHAS,RAD,TAX
     3.Numeric- LON,LAT,MEDV,CRIM,ZN,INDUS,NOX,RM,AGE,DIS,PTRATIO
     4.Number of columns:16
     5.Number of observation:506
     
```{r}
# Summary of dataset
summary(dataset)
```

**OBSERVARTION:**
        The house price ranges between 5000 and 50000 dollars.
        The Longitude and Latitude lies on -71 and 42 coordinate points respectively in Boston.
        
```{r}
# Checking for NA values.
colSums(is.na(dataset))
```

```{r}
# Checking for Empty values.
colSums(dataset=='')
```

**OBSERVARTION:**
        There is no presence NA and empty values in Boston dataset.
        
```{r}
# Checking for outliers if any
j=c(3,4)
for (i in j)
{
x=dataset[,i]
boxplot(x,col='grey',
     xlab=colnames(dataset[i]),
     main=colnames(dataset[i]))
}
```

```{r}
# Removing outliers
outliers <- boxplot(dataset$LON, plot=FALSE)$out 
print(outliers)

outliers1 <- boxplot(dataset$LAT, plot=FALSE)$out 
print(outliers1)
```

```{r}
# Outliers spotted in dataset
dataset[which(dataset$LON %in% outliers),]
dataset <- dataset[-which(dataset$LON %in% outliers),]
```

```{r}
dataset[which(dataset$LAT %in% outliers1),]
dataset <- dataset[-which(dataset$LAT %in% outliers1),]
```

**OBSERVARTION:**
        Outliers are spotted out using boxplot. Here we notice that longitude and latitude values have outliers. So the outliers are firstly spotted and removed from the dataset.  

### QUESTION 2.MEDV is the output /target variable i.e price of the house to be predicted.

```{r}
summary(dataset$MEDV)
```

```{r}
par(mfrow=c(2,3))

plot(dataset$LON, dataset$MEDV, main = "Longitude", xlab="", ylab="", cex.lab="1.5", col="#1e3799", pch=20, cex= 0.5 ) 
plot(dataset$LAT, dataset$MEDV, main = "Latitude", xlab="", ylab="", cex.lab="1.5", col="#628ca6", pch=20, cex= 0.5)
plot(dataset$CRIM, dataset$MEDV, main = "Crime Rate", xlab="", ylab="", cex.lab="1.5", col="#e58e26", pch=20, cex= 0.5)
plot(dataset$ZN, dataset$MEDV, main = "Residential Zone", xlab="", ylab="", cex.lab="1.5", col="#628ca6", pch=20, cex= 0.5)
plot(dataset$INDUS, dataset$MEDV, main = "Industry Area", xlab="", ylab="", cex.lab="1.5", col="#e58e26", pch=20, cex= 0.5)
plot(dataset$TAX, dataset$MEDV, main = "Tax per $10,000", xlab="", ylab="", cex.lab="1.5", col="#1e3799", pch=20, cex= 0.5)
```
```{r}
par(mfrow=c(2,3))

plot(dataset$NOX, dataset$MEDV, main = "Nitrous Oxides", xlab="", ylab="", cex.lab="1.5", col="#84817a", pch=20, cex= 0.5 ) 
plot(dataset$RM, dataset$MEDV, main = "Number of Rooms", xlab="", ylab="", cex.lab="1.5", col="#cd6133", pch=20, cex= 0.5 ) 
plot(dataset$CRIM, dataset$MEDV, main = "Crime Rate", xlab="", ylab="", cex.lab="1.5", col="#b33939", pch=20, cex= 0.5 ) 
plot(dataset$AGE, dataset$MEDV, main = "Built before 1940 (Age)", xlab="", ylab="", cex.lab="1.5", col="#b33939", pch=20, cex= 0.5 ) 
plot(dataset$DIS, dataset$MEDV, main = "Employment", xlab="", ylab="", cex.lab="1.5", col="#cd6133", pch=20, cex= 0.5 ) 
plot(dataset$RAD, dataset$MEDV, main = "Highways", xlab="", ylab="", cex.lab="1.5", col="#84817a", pch=20, cex= 0.5 ) 
```

**OBSERVATIONS:**
       1.There is an observation that when house prices go down there is a significant increase in crime rate.
       2.There is presence of nitous oxide in air which cause air pollution.And its normally distributed in all range of house price regions.
       3.There is higher probability of employment opportunity when the housing price significantly increases.


### QUESTION 3. Using the plot commands, plot the latitude and longitude of each of our census tracts.
```{r}
plot(dataset$LON, dataset$LAT, main = "Longitude and Latitude of census tracts",  xlab="Longitude", ylab="Latitude", col="#bbbbbb", pch=20, cex= 1.0)
```
```{r}
library(plotly)
df <- read.csv('boston.csv')

# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

fig <- plot_geo(df, lat = ~df$LAT, lon = ~df$LON)
fig <- fig %>% add_markers(
    text = ~paste(df$LAT, df$LON, df$Town, sep = "<br />"),
    color = ~df$TOWN, symbol = I("square"), size = I(8),hoverinfo = "Text"
  )
fig <- fig %>% colorbar(title = "TOWN")
fig <- fig %>% layout(title = 'Longitude and Latitude of census tracts<br />(Boston)', geo = g,hovermode = TRUE)

fig
```

**OBSERVATIONS:**
           It shows that there is no linearity and moderate relationship between latitude&longitude over house locations.
           
           

### QUESTION 4. Show all the points that lie along the Charles River in a blue colour.
```{r}
# https://colorhunt.co/

plot(dataset$LON, dataset$LAT, main = "Charles River in a blue colour",  xlab="Longitude", ylab="Latitude", col="#bbbbbb", pch=20, cex= 1)
points(dataset$LON[dataset$CHAS==1], dataset$LAT[dataset$CHAS==1], col="#005fba", pch=16)
```

**OBSERVATIONS:**
           Blue colored highlighted house locations were generally lies along the Charles River.

### QUESTION 5.Apply Linear Regression by plotting the relationship between latitude and house prices and the longitude and the house prices.
```{r}
library(plotly)
fig <- plot_ly(data = dataset, x = ~dataset$LAT, y = ~dataset$MEDV, marker = list(size = 10,color = '#822659'))
fig <- fig %>% layout(title = 'RELATIONSHIP LATITUDE AND HOUSE PRICE',yaxis = list(title = "House Price",zeroline = TRUE),xaxis = list(title = "Latitude",zeroline = TRUE))

fig
```


```{r}
fig <- plot_ly(data = dataset, x = ~dataset$LON, y = ~dataset$MEDV, marker = list(size = 10,color = '#005AFF'))
fig <- fig %>% layout(title = 'RELATIONSHIP LONGITUDE AND HOUSE PRICE ',yaxis = list(title = "House Price",zeroline = TRUE),xaxis = list(title = "Longitude",zeroline = TRUE))

fig
```

**OBSERVATIONS:**
           Above scatter plot shows that they are randomly distributed and looks like there is no relationship between them.
           
```{r}
set.seed(35)
spts<-sample(1:nrow(dataset),0.7*nrow(dataset))#70% Training Data
train<-dataset[spts,]
test<-dataset[-spts,]
```


```{r}
#linear model
linear_model = lm(MEDV ~ LAT + LON, data=dataset)
summary(linear_model)
```

**OBSERVATIONS:**
        - The estimated regression model is:
          Housing prices =  -5335.14+ 12.85(LAT) + -67.76(LON)
        - Standard error shows the approximate variation of the response variable with respect to the predictor variable.
        - The average house price is -5335.14 when predictor variables are zero.
        - p-value is 2.2e-16 < 0.05,hence there is no relationship between response and predictor variable.
        - 45.13 on 2 and 460 DF, this value is far from 1, hence the model isn't significant.
        - Multiple R-squared:  0.164, 16.4 % well fits the model.
         

```{r}
plot(dataset$LON, dataset$LAT, main = "Longitude and Latitude of census tracts",  xlab="Longitude", ylab="Latitude", col="black", pch=20, cex= 1)
points(dataset$LON[dataset$MEDV >= 21.2], dataset$LAT[dataset$MEDV >= 21.2], col="#005fba", pch=16)
points(dataset$LON[linear_model$fitted.values >= 21.2],dataset$LAT[linear_model$fitted.values >= 21.2],col='red',pch=16)
```

```{r}
pred <- predict(linear_model,test)
compare <- cbind(actual=test$MEDV,pred)
head(compare)
```
```{r}
#Mean Square Error
mean((test$MEDV-pred)^2)
```
```{r}
#Root mean Square Eroor
library(Metrics)
rmse(pred, test$MEDV)
```

**OBSERVATION**
            Therefore, we can observe that our model isn't good fit.

### QUESTION 6.Apply Regression Tree to the problem and draw conclusions from it.
```{r}
library(rpart)
library(rpart.plot)
```

```{r}
tree = rpart(MEDV ~ LAT + LON + CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX + PTRATIO, data=dataset)
prp(tree)
```
```{r}
fittedvalues=predict(tree)
plot(dataset$LON, dataset$LAT, main = "Longitude and Latitude of census tracts",  xlab="Longitude", ylab="Latitude", col="black", pch=20, cex= 1)
points(dataset$LON[dataset$MEDV >= 21.2], dataset$LAT[dataset$MEDV >= 21.2], col="#005fba", pch=16)
points(dataset$LON[fittedvalues >= 21.2],dataset$LAT[fittedvalues >= 21.2],col='red',pch=16)
```



```{r}
# Make predictions on the test data
predictions <- predict(tree,test)
compare1 <- cbind(actual=test$MEDV,predictions)
head(compare1)
```
```{r}
rmse(predictions, test$MEDV)
```

**OBSERVATIONS:**
      We can observe CART model is better than Linear Regression model both by visually and by RMSE.