---
title: "P7"
author: "SANDRA MARIA JOSEPH"
date: "14/03/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Lab#7 Demonstrate the program using Regression Tree**

# IMPORTING DATASET
```{r}
#importing boston dataset

data=read.csv("E:/MDS_SEM2/New folder/R_LAB/PROGRAM7/boston.csv")
head(data)
```
```{r}
#finding the dimension of the dataset
dim(data)
```
```{r}
#finding the Column names
colnames(data)
```
The boston dataset contains 506 observations in 16 columns.
The columns are:

  LON and LAT are the longitude and latitude of the center of the census tract.
  MEDV is the median value of owner-occupied homes, measured in thousands of dollars.
  CRIM is the per capita crime rate.
  ZN is related to how much of the land is zoned for large residential properties.
  INDUS is the proportion of the area used for industry.
  CHAS is 1 if a census tract is next to the Charles River else 0
  NOX is the concentration of nitrous oxides in the air, a measure of air pollution.
  RM is the average number of rooms per dwelling.
  AGE is the proportion of owner-occupied units built before 1940.
  DIS is a measure of how far the tract is from centres of employment in Boston.
  RAD is a measure of closeness to important highways.
  TAX is the property tax per $10,000 of value.
  PTRATIO is the pupil to teacher ratio by town

# DATA PREPROCESSING  
```{r}
#checking for NAN values
colSums(is.na(data))
```
```{r}
#checking NULL values
is.null(data)
```
```{r}
#checking EMPTY values
colSums(data=='')
```
## The dataset is clean.It is free from nan and null values.

```{r}
#removing Town and Tract

data_new<-subset(data,select=-c(TOWN,TRACT))
head(data_new)

```
# EXPLORATORY DATA ANALYSIS

```{r}
#structure of the dataset
str(data_new)
```
### CHAS, RAD and Tax are of Integer type while all other variables are of numerical data type.

```{r}
#sample descriptives
library(psych)
describe(data_new)
```
```{r}
summary(data_new)
```
```{r}
j=c(1,2,4,5,6,7,8,9,10,11,12,13,14)
for (i in j)
{
  library(car)
X=data_new[,i]
Y=data_new$MEDV
scatterplot(X,
            Y,
            xlab =colnames(data_new[i]),
            ylab=colnames(data_new[3]),
            boxplots = " ", # Disable boxplots
            smooth = FALSE, # Removes smooth estimate
            regLine = TRUE, # add regression line
            grid = TRUE,
            main=colnames(data_new[i])
)
}
```
# ZN, RM, DIS have a positive trend with MEDV.

```{r}
#Checking Normality of Response Variable

library(e1071)
plot(density(data_new$MEDV), 
     main = "Price of the house",
     ylab="Frequency",
     sub=paste("Skewness",round(e1071::skewness(data_new$MEDV),2)))
polygon(density(data_new$MEDV), col='#FFC9A9')
```
```{r}
j=c(1,2,4,5,6,7,8,9,10,11,12,13,14)
for (i in j)
{
x=data_new[,i]
hist(x,col='yellow',
     xlab=colnames(data_new[i]),
     main=colnames(data_new[i]))
}
```
```{r}
j=c(1,2,4,5,6,7,8,9,10,11,12,13,14)
for (i in j)
{
x=data_new[,i]
c=density(x)
plot(c,col='black',
     xlab=colnames(data_new[i]),
     main=colnames(data_new[i]),
     sub=paste("Skewness",round(e1071::skewness(x),2)))
polygon(c, col = "magenta")

}
```
*3. Using the plot commands, plot the latitude and longitude of each of our census tracts.*
```{r}
plot(x=data_new$LON,
     y=data_new$LAT,
     xlab="LONGITUDE",
     ylab="LATITUDE",
     main="Latitude VS Longitude",
     pch=1,
    col='green')

```
```{r}
library("ggplot2")
library("gghighlight")
#data_new$CHAS<-as.factor(data_new$CHAS)
ggplot(data_new,aes(x=LON,y=LAT,color=CHAS))+geom_point()+labs(title="Latitude & Longitude of Census Track")
```
*4. Show all the points that lie along the Charles River in a blue colour.*
```{r}
plot(x=data_new$LON,
     y=data_new$LAT,
     xlab="LONGITUDE",
     ylab="LATITUDE",
     main="Latitude VS Longitude",
     col='red',
     pch='+')
points(data_new$LON[data_new$CHAS==1], data_new$LAT[data_new$CHAS==1], col="blue", pch=18)
```
*5.Apply Linear Regression by plotting the relationship between latitude and house prices and the longitude and the house prices.*

```{r}
ggplot(data=data_new,aes(x=LON,y=MEDV))+stat_smooth(method="lm")+labs(title="Longitude & House Prices",x="Longitude",y="House Price")+geom_point(aes(color=factor(CHAS)))
```
```{r}
ggplot(data=data_new,aes(x=LAT,y=MEDV))+stat_smooth(method="lm")+labs(title="Latitude & House Prices",x="Latitude",y="House Price")+geom_point(aes(color=factor(CHAS)))
```
```{r}
#correlation between Latitude and MEDV
cor(data_new$LAT,data_new$MEDV)
```
```{r}
#correlation between LON and MEDV
cor(data_new$LON,data_new$MEDV)
```
# Splitting the dataset

```{r}
#splitting the dataset

set.seed(35)
spts<-sample(1:nrow(data_new),0.7*nrow(data_new))#70% Training Data
train<-data_new[spts,]
test<-data_new[-spts,]
```

```{r}
#linear regression model
lrm <- lm(MEDV~LAT+LON,data=train)
lrm
```
```{r}
summary(lrm)
```
# The regression equation is MEDV= -2722.7072+0.3813*LAT -38.4097*LON.
# The value of adjusted Rˆ2 is 0.1043 which means the model can explaIN 10.43% variability in MEDV.

```{r}
pred <- predict(lrm,test)
compare <- cbind(actual=test$MEDV,pred)
head(compare)
```
```{r}
#Mean Square Error

mean((test$MEDV-pred)^2)
```
```{r}
#Root mean Square Eroor

RMSE = sqrt(mean((test$MEDV-pred)^2))
RMSE# calculate accuracy
```
The obtained RMSE value is 9.069179.

__________________________________________________________________________________________________________ 
  *6.Apply Regression Tree to the problem and draw conclusions from it.*
```{r}
library(caret)
```


```{r}
# Fit the model on the training set
set.seed(123)

model <- train(
  MEDV ~LAT+LON,
  data = train, 
  method = "rpart",
  trControl = trainControl("cv", number = 10),   #to set up 10-fold cross validation
  tuneLength = 10                                #to specify the number of possible cp values to evaluate.                                                   Default value is 3, here we use 10.
  )

# Plot model error vs different values of
# cp (complexity parameter)
plot(model)

# Print the best tuning parameter cp that
# minimize the model RMSE
model$bestTune
```

```{r}
# Plot the final tree model
par(xpd = NA) # Avoid clipping the text in some device
plot(model$finalModel)
text(model$finalModel, digits = 3)

```

```{r}
# Decision rules in the model
model$finalModel
```


```{r}
# Make predictions on the test data
predictions <- predict(model,test)
head(predictions)
```


```{r}
# Compute the prediction error RMSE
RMSE(predictions, test$MEDV)
```

```{r}
library(rpart)
library(rpart.plot)
```


```{r}
rtree.fit <- rpart(MEDV ~LAT + LON + CRIM + ZN + INDUS + CHAS + NOX + RM + AGE + DIS + RAD + TAX+PTRATIO, 
                  data=train,
                  method="anova",  #for regression tree
                  control=rpart.control(minsplit=30,cp=0.001))#optional parameters for controlling tree                                                                   growth.
```

  
```{r}
#displays table of fits across cp (complexity parameter) values

printcp(rtree.fit) # display the results 
```
```{r}
#plot approximate R-squared and relative error for different splits
rsq.rpart(rtree.fit) #produces 2 plots
```
```{r}
plotcp(rtree.fit) # visualize cross-validation results 
```
# A good choice of cp for pruning is often the leftmost value for which the mean lies below the horizontal line.

```{r}
summary(rtree.fit) # detailed summary of splits
```
```{r}
library(rpart.plot)
rpart.plot(rtree.fit)
```


```{r}
plot(rtree.fit, uniform=TRUE, 
    main="Regression Tree for Median Home Value")
text(rtree.fit, use.n=TRUE, all=TRUE, cex=.8)
```
```{r}
prp(rtree.fit)
```
```{r}
# prune the tree based on minimim xerror
pruned.rtree.fit<- prune(rtree.fit, cp= rtree.fit$cptable[which.min(rtree.fit$cptable[,"xerror"]),"CP"])

# plot the pruned tree using prp() in the rpart.plot package 
prp(pruned.rtree.fit, main="Pruned Regression Tree for Median Home Value")
```
```{r}
# prune the tree based on 1 SE error 
pruned2.rtree.fit<- prune(rtree.fit, cp=.01)

# plot the pruned tree using prp() in the rpart.plot package
prp(pruned2.rtree.fit, main="Pruned Regression Tree for Median Home Value")

```
```{r}
# Make predictions on the test data
predictions <- predict(rtree.fit,test)
head(predictions)

# Compute the prediction error RMSE
RMSE(predictions, test$MEDV)
```


```{r}
# Make predictions on the test data
predictions <- predict(pruned.rtree.fit,test)
head(predictions)

# Compute the prediction error RMSE
RMSE(predictions, test$MEDV)
```



```{r}
# Make predictions on the test data
predictions <- predict(pruned2.rtree.fit,test)
head(predictions)

# Compute the prediction error RMSE
RMSE(predictions, test$MEDV)
```



