---
title: "LAB 2"
author: "2048015 - MANOJ KUMAR"
date: "29/01/2021"
output: word_document
---


### 1.Import the LungCapData into R and attach it.

```{r}
## Library for importing/reading Excel dataset.
library(readxl)

dataframe <- read_excel("Lungcap.xlsx")
dataframe
```

### 2. Find the class and typeof Age and Height.

```{r}
## Age

sprintf('The class of Age is: %s', class(dataframe$Age))
sprintf('The type of Age is: %s', typeof(dataframe$Age))
```


```{r}
## Height

sprintf('The class of Height is: %s', class(dataframe$Height))
sprintf('The type of Height is: %s', typeof(dataframe$Height))
```

### 3. Find the summary of the dataset.

```{r}
## Summary function describe entire dataset.

summary(dataframe)
```

### 4.Find the strength of the relationship between Age and Height (Hint:Using Correlation).

```{r}
## Correlation function

cor(dataframe$Age,dataframe$Height)

#The correlation coefficient is a number between 1 and -1
```

### 5.Perform a scatterplot for the above (Qn.4).

```{r}
## Case 1

x <- dataframe$Age
y <- dataframe$Height

plot(x, y, 
     main = "Correlation between Age and Height", 
     xlab = "Age", 
     ylab = "Height",     
     cex.lab = "1",
     col = "Brown",
     pch = 15
     )
```


```{r}
## Case 2

# Importing ggplot for plotting 
library(ggplot2)

x <- dataframe$Age
y <- dataframe$Height

ggplot(dataframe, aes(x, y)) + geom_point()
```

### 6. Add a title to the plot and Label  x-axis and y-axis.

```{r}
plot(x, y, 
     main = "Correlation between Age and Height", 
     xlab = "Age",
     ylab = "Height",     
     cex.lab = "1",
     col = "Blue",
     pch = 10
     )
```

### 7.Rotate the values on the y-axis and also change the limits of x-axis and y-axis.

```{r}
## Case 1
plot(x, y, 
     main = "Correlation between Age and Height", 
     xlab = "Age",
     ylab = "Height",
     xlim = c(1, 25),
     ylim = c(30, 90),
     cex.lab = "1",
     col = "Black",
     pch = 10
     )
```
```{r}

plot = ggplot(dataframe, aes(x, y)) + geom_point() 
plot + theme(axis.text.y = element_text (angle = 60, hjust = 1))
```


### 8.  Remove the axes of the plot and relabel these axes.

```{r}
plot(x, y, 
     xlab = "Age", ylab = "Height",
     xlim = c(1,25), ylim = c(40,85),
     axes = FALSE, pch = 8
     )
```

### 9.Change the size of the plotting characters.

```{r}
plot(x, y, 
     xlab = "Height", ylab = "Age" ,
     main = "Age Vs Height",
     pch = 15, cex= 0.2
     )
```

### 10.Change the color of the characters to red.

```{r}
plot(x, y, 
     xlab = "Height", ylab = "Age" ,
     main = "Age Vs Height",
     col = "Red", pch = 15, cex= 0.2
     )
```

### 11.Draw the regression line to the plot predicting height using age.

```{r}
# Importing ggplot for plotting 

library(ggplot2)

ggplot(dataframe, aes(x, y)) + geom_point() + geom_smooth(method = lm, 
                                                          color = "red")
```

### 12.Change the color of the line to blue and the width of the line.

```{r}

ggplot(dataframe, aes(x, y)) + geom_point() + geom_smooth(method = lm, 
                                                          color = "blue",
                                                          size = 0.2 )
```

### 13.Change the font of the plotting characters.

```{r}

plot + geom_smooth (method = lm, color = "blue", size = 0.3) + geom_point(shape=15)

```

### 14.Change the color of the plotting characters.

```{r}

plot + geom_smooth(method = lm, 
                   color = "blue",
                   size = 0.3) + geom_point (shape = 16, color = "#e28743")
``` 

### 15.Identify gender on the plot for the age male and female.

```{r}
plot + geom_smooth(method = lm, 
                   color = "blue",
                   size = 0.3) + geom_point (shape = 16,aes(color=Gender))
```

