---
title: "LAB 2"
author: "2048015 - MANOJ KUMAR"
date: "25/01/2021"
output: word_document
---


### 1. Create a vector different data types(Logical, Numeric, Integer, Complex, Character) and display their class and typeof each datatype.

```{r}
# Logical Datatype

Vector_Logical <- c(TRUE, FALSE , TRUE, FALSE)
sprintf('The class of Vector_Logical is: %s', class(Vector_Logical))
sprintf('The type of Vector_Logical is: %s', typeof(Vector_Logical))
```


```{r}
# Numeric Datatype

Vector_Numeric <- c(2048015, 1948015, 1848015)
sprintf('The class of Vector_Numeric is: %s', class(Vector_Numeric))
sprintf('The type of Vector_Numeric is: %s', typeof(Vector_Numeric))
```

```{r}
# Integer Datatype

Vector_Integer <- c(75L, 85L)
sprintf('The class of Vector_Integer is: %s', class(Vector_Integer))
sprintf('The type of Vector_Integer is: %s', typeof(Vector_Integer))
```

```{r}
# Complex Datatype

Vector_Complex <- c(2 + 4i, 2 + 3i)
sprintf('The class of Vector_Complex is: %s', class(Vector_Complex))
sprintf('The type of Vector_Complex is: %s', typeof(Vector_Complex))
```

```{r}
#Character Datatype

Vector_Char <- c('Manoj', "Kumar", "MDS")
sprintf('The class of Vector_Char is: %s', class(Vector_Char))
sprintf('The type of Vector_Char is: %s', typeof(Vector_Char))
```

### 2. Get and print the current working directory.

```{r}

sprintf('The current directory is: %s', getwd())
```

```{r}
# To set the directory.

setwd("/Users/manojkumarr/Downloads/Other/Christ University/SEM 2/R/Lab")
sprintf('The new directory is: %s', getwd())
```

### 3. Create this file using windows notepad by copying and pasting this data. Save the file as student.csv

S.No ,Sname, Degree, Total.marks, Grade
1,Andrew,UG,435,B
2,Babita,UG,210,D
3,Cathy,UG,459,A
4,Dominic,UG,542,A
5,Elsa,PG,520,B
6,Franko,PG,320,C
7,Gorang,UG,205,D 
8,Harsha,PG,325,C 

```{r}
## student.csv file created.
```

### 4.Save this file in the current working directory.

```{r}
## File Saved in the current working directory 
```


### 5.Read the csv file in your current working directory.

```{r}
student_details<-read.csv("student.csv")
student_details
```

### 6.Check whether your  CSV file is a dataframe and also check the number of rows and columns.

```{r}
sprintf('The class of df is: %s', class(student_details))
```
```{r}
# nrow(df) gives the number of rows in df.

sprintf('Number of rows in df is: %s', nrow(student_details))
```
```{r}
# ncol(df) gives the number of columns in df.

sprintf('Number of columns in df is: %s', ncol(student_details))
```

### 7. Apply all the functions sum(),mean(),sqrt()related to dataframe.

```{r}
str(student_details)
```

**Total marks is the only column that is numerical. So we will be performing the functions in that particular column only**
```{r}
Total_Marks <- student_details$Total.marks
Total_Marks
```

```{r}
# SUM

sprintf('Sum of Total marks is: %s', sum(Total_Marks))
```

```{r}
# MEAN

sprintf('Mean of Total marks is: %s', mean(Total_Marks))
```

```{r}
# Square root of Total Marks for each student is:-

sprintf('Square root of Total marks is: %s',sqrt(Total_Marks))
```

### 8.Get the highest marks from the data frame.

```{r}
HighestMark<-max(student_details$Total.marks)
sprintf('The Highest Marks is: %s', HighestMark)
```

### 9. Get the details of the person with highest marks.

```{r}
# Case 1

index=which.max(student_details$Total.marks)
student_details[index,]
```


```{r}
# Case 2

HighestScorer = subset(student_details, Total.marks==max(student_details$Total.marks))
HighestScorer
```


```{r}
# Case 3

student_details[student_details$Total.marks == max(student_details$Total.marks),]
```

### 10. Get all the students in UG degree whose marks is greater than 300.

```{r}
Students_list1 <- student_details[(student_details$Degree == 'UG') & (student_details$Total.marks > 300), ]
Students_list1
```

### 11. Add one more vector  Date_ of_ Joining(DOJ)  to the already existing dataframe.
```{r}
library(lubridate)
```

```{r}
doj <- c("20160505","20160507","20170509","20180517","20290621","20200625","20200627", "20200611")

student_details$Date_of_Joining<-as.Date(doj, format = "%Y%m%d")
student_details
```

### 12. Get the details of the students who have joined after 2017.

```{r}
Student_list2 <- student_details[year(student_details$Date_of_Joining) > 2017, ]
Student_list2
```

### 13. Write the filtered data into a new file.

```{r}
write.csv(Student_list2, "/Users/manojkumarr/Downloads/Other/Christ University/SEM 2/R/Lab/result.csv")
```