---
title: "R-Laboratory 6"
author: "MANOJ KUMAR - 2048015"
date: "26/02/2021"
output: word_document
---

## 1.Load mtcars dataset.

```{r}
# Loading mtcars dataset.
data(mtcars)
mtcars
```

*Insight *

        - mtcars dataset contains 32 obs. of  11 variables.
        - Notably all the variables are Numerical.
        
## 2.install ridge and glmnet packages.

```{r}
#install.packages(ridge)
#install.packages(glmnet)
```
```{r}
library(ridge)      # Linear and logistic ridge regression functions.
```


```{r}
library(Matrix)
```


```{r}
library(glmnet)     # Lasso and Elastic-Net Regularized Generalized Linear Models
```

## 3.Perform the exploratory data analysis.

```{r}
# Pre-processing EDA
df = mtcars

str(df)
```

*Insight *

        - Totally, 32 observations of  11 variables.
        - All the 11 features are numerical datatypes. 

```{r}
# Summary

summary(df)
```

*Insight *

        - Minimum gear level is 3 and maximum gear level is 5
        - Minimum cylinders capacity is 4 and maximum capacity is 8
        - Minimum horsepower is 52 and maximum value is 335
        - Number of carburetors minimum is 1 and maximum of 8

```{r}
# Checking for missing values.

colSums(is.na(df))
```

```{r}
#Checking for Empty Values

colSums(df=='')
```

```{r}
#Checking for Duplicate values

library(tidyverse)
```
```{r}
duplicated(df)    # Row
```


*Insight *

        - The dataset is clean, there are no missing, Empty values or duplicated record.

```{r}
# Checking Normality of Response Variable

# Using this method, we obtain predictions from the model, as well as decision values from the binary classifiers.
library(e1071)

# plot() – Visualizing data, support vectors and decision boundaries, if provided.
plot(density(df$mpg), 
     main = "Millage Density Plot", 
     ylab="Frequency",  
     sub=paste("Skewness",round(e1071::skewness(df$mpg),2))
     )

polygon(density(df$mpg), col='#079992')
```

*Insight *

        - Slightly Right Skewked, which implies most of the values are positive in nature.
        - Mode > Median > Mean values 
        - Skewness  value is > 0, so data values are less than mean
        - This chart is a variation of a Histogram that uses kernel smoothing to plot values, allowing for smoother distributions by smoothing out the noise.


```{r}
#Correlation Heatmap

library(ggplot2)        # Makes it simple to create complex plots from data in a data frame.
```


```{r}
library(reshape2)       
# Long- to wide-format data: the cast functions  
# Wide- to long-format data: the melt function
```


```{r}
cormat <- round(cor(df),2)
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

*Insight *

        - Darker shades denotes less correlationship indications.
        - Lighter shades denotes High correlationship with each variables.
        - It is evident that most of the variables possess a high correlation with each other, thus we can assume multicollinearity is present.
        - We need to checking for outliers in highly positive correlated values.

```{r}
#Checking for outliers in highly positive correlated values with Mileage

par(mfrow=c(2,3))

boxplot(df$drat, main = "Rear axle ratio")
boxplot(df$qsec, main = "1/4 mile time")
boxplot(df$gear, main = "Number of Forward Gears")
boxplot(df$hp,   main = "Gross horsepower")
boxplot(df$wt, main = "Weight (lb/1000)")
boxplot(df$disp,   main = "Displacement (cu.in.)")

```

*Insight *

        - Clearly there is an outlier found in qsec(1/4 mile time).
        - Also, there is an outlier found in Gross horsepower.
        - Clearly huge outlier were found in Weight (lb/1000).


```{r}
# Splitting the data
set.seed(57)   # Starting number used to generate a sequence of random numbers  

# sample() allows you to take a random sample of elements from a dataset or a vector, either with or without replacement.
trainingRow <- sample(1:nrow(df), 0.7*nrow(df))

trainset <- df[trainingRow,]
testset <- df[-trainingRow,]

lrm <- lm(trainset$mpg~., data=trainset)

summary(lrm)
```
*Insight *

        - Sequence of random numbers is seeded and taken a random sample of elements from a dataset for the trainset and testset.
        - Linear regression model is created using the trainset.
        - Estimate, S.D, t-value, Residuals information are summarize from created model.
        - Residual standard error: 3.006 on 11 degrees of freedom
        - RSE measures the average of the squares of the errors. Lower values (closer to zero) indicate better fit.


```{r}
# Variance Inflation Factor 
vif(lrm)
```


*Insight *
        
        - VIF measures how much the variance of a regression coefficient is inflated due to multicollinearity in the model.
        - Most of the values are above 5, there is strong multi-collinearity present.
        - Between 5 to 10 multicollinearity and we should remove those features.
      
```{r}
# Multi linear reg 
MLR_pred <- predict(lrm,testset)
compare <- cbind(actual=testset$mpg, MLR_pred)
compare
```

```{r}
mean (apply(compare, 1, min)/apply(compare, 1, max))
```

```{r}
RMSE = sqrt(mean((testset$mpg-MLR_pred)^2))
RMSE# calculate accuracy
```

*Insight *

        - Accuracy is only 86%, which is not very efficient.

## 4.Choose optimum lamba value.

```{r}
#Building initial model
X = model.matrix(mpg~. , mtcars)[,-1]   # Matrix with Response variables or Dependent variable with others mpg~.
Y = mtcars$mpg
```


```{r}
#Creating a sequence with an interval of -0.12
lambda_seq = 10^seq(3, -2, by = -.12)

# Using cross validation glmnet
ridge_model1 = cv.glmnet(X[trainingRow,], Y[trainingRow],alpha = 0, type.measure = "mse", lambda = lambda_seq, nfolds = 5)

# Best lambda value
best_lam =  ridge_model1$lambda.min
best_lam
```

```{r}
plot(ridge_model1)
```

*Insight *

        - Optimum lamba value choosed and plotted. (lambda_seq and 2.290868)
        - nfolds = 5 Cross validation
        
## 5.Extract the model using k-cross validation.

```{r}
best_fit <- ridge_model1$glmnet.fit
head(best_fit)
```

*Insight *

        - Cross validation fit explained

## 6.Build the final model and interpret.

```{r}

# linearRidge

linRidgeMod = linearRidge(trainset$mpg ~ ., data = trainset)
predicted = predict(linRidgeMod, testset) # predict on test data
compare1 = cbind (actual=testset$mpg, predicted)

mean (apply(compare1, 1, min)/apply(compare1, 1, max))

RMSE = sqrt(mean((testset$mpg-predicted)^2))
RMSE

summary(linRidgeMod)
```

*Creating another model with only significant values.*

```{r}
# wt, gear, crab

linRidgeMod = linearRidge(trainset$mpg ~ ., data = trainset[, c(6,10,11)])
predicted1 = predict(linRidgeMod, testset) # predict on test data
compare2 = cbind (actual=testset$mpg, predicted1)

summary(linRidgeMod)

mean (apply(compare2, 1, min)/apply(compare2, 1, max))

RMSE = sqrt(mean((testset$mpg-predicted1)^2))
RMSE
```

*Insight *

        - The accuracy has increased from 85% to 94%.
        - Root-Mean-Square Error has decreased from 3.242 to 1.389


