---
title: "R-Laboratory 6"
author: "MANOJ KUMAR - 2048015"
date: "26/02/2021"
output: word_document
---

## 1.Load mtcars dataset.

```{r}
# Loading mtcars dataset.
data(mtcars)
mtcars
```

## 2.install ridge and glmnet packages.

```{r}
#install.packages(ridge)
#install.packages(glmnet)
```
```{r}
library(ridge)      # Linear and logistic ridge regression functions.
library(Matrix)
library(glmnet)     # Lasso and Elastic-Net Regularized Generalized Linear Models
```

## 3.Perform the exploratory data analysis.

```{r}
# Pre-processing EDA
df = mtcars

str(df)
```

*Insight *

        - Totally, 32 observations of  11 variables.
        - All the 11 features are numerical datatypes. 

```{r}
# Summary

summary(df)
```


```{r}
# Checking for missing values.

colSums(is.na(df))
```


```{r}
#Checking for Empty Values

colSums(df=='')
```

```{r}
#Checking for Duplicate values

library(tidyverse)
```
```{r}
duplicated(df)
```


*Insight *

        - The dataset is clean, there are no missing, Empty values or duplicated record.

```{r}
# Checking Normality of Response Variable

# Using this method, we obtain predictions from the model, as well as decision values from the binary classifiers.
library(e1071)


# plot() – Visualizing data, support vectors and decision boundaries, if provided.
plot(density(df$mpg), 
     main = "Milage Density Plot", 
     ylab="Frequency",  
     sub=paste("Skewness",round(e1071::skewness(df$mpg),2))
     )

polygon(density(df$mpg), col='#079992')
```

*Insight *

        - Slightly Right Skweked, which implies most of the values are posititve in nature. 

```{r}
#Correlation Heatmap

library(ggplot2)
library(reshape2)
```


```{r}
cormat <- round(cor(df),2)
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```

*Insight *

        - Darker shades denotes less correlationship indications.
        - Lighter shades denotes High correlationship with each variables.
        - It is evident that most of the variables possess a high correlation with each other, thus we can assume multicollinearity is present.

```{r}
#Checking for outliers in highly positive correlated values with Mileage

par(mfrow=c(2,3))

boxplot(df$drat, main = "Rear axle ratio")
boxplot(df$qsec, main = "1/4 mile time")
boxplot(df$gear, main = "Number of Forward Gears")
boxplot(df$hp,   main = "Gross horsepower")
boxplot(df$cyl, main = "Number of cylinders")
boxplot(df$disp,   main = "Displacement (cu.in.)")

```

*Insight *

        - There is an outlier found in qsec(1/4 mile time) and Gross horsepower.

```{r}
#Building initial model
X = model.matrix(mpg~. , mtcars)[,-1]
Y = mtcars$mpg

#Splitting the data
set.seed(57)

trainingRow <- sample(1:nrow(df), 0.7*nrow(df))
trainset <- df[trainingRow,]
testset <- df[-trainingRow,]

lrm <- lm(trainset$mpg~., data=trainset)

summary(lrm)

```


```{r}
library(car)

vif(lrm)
```


*Insight *

        - All the values are above 5, there is strong multi-collinearity present.
      
```{r}
MLR_pred <- predict(lrm,testset)
compare <- cbind(actual=testset$mpg, MLR_pred)
compare
```

```{r}
mean (apply(compare, 1, min)/apply(compare, 1, max))
```

```{r}
RMSE = sqrt(mean((testset$mpg-MLR_pred)^2))
RMSE# calculate accuracy

```

*Insight *

        - Accuracy is only 81%, which is not very efficient.

## 4.Choose optimum lamba value.

```{r}
#Creating a sequence with an interval of -0.12
lambda_seq = 10^seq(3, -2, by = -.12)

# Using cross validation glmnet
ridge_model1 = cv.glmnet(X[trainingRow,], Y[trainingRow],alpha = 0, type.measure = "mse", lambda = lambda_seq, nfolds = 5)

# Best lambda value
best_lam =  ridge_model1$lambda.min
best_lam
```

```{r}
plot(ridge_model1)
```

*Insight *

        - Optimum lamba value choosed and plotted.
        
## 5.Extract the model using k-cross validation.

```{r}
best_fit <- ridge_model1$glmnet.fit
head(best_fit)
```

## 6.Build the final model and interpret.

```{r}
linRidgeMod = linearRidge(trainset$mpg ~ ., data = trainset)
predicted = predict(linRidgeMod, testset) # predict on test data
compare1 = cbind (actual=testset$mpg, predicted)

mean (apply(compare1, 1, min)/apply(compare1, 1, max))

summary(linRidgeMod)
```

*Creating another model with only significant values.*

```{r}
linRidgeMod = linearRidge(trainset$mpg ~ ., data = trainset[, c(6,10,11)])
predicted1 = predict(linRidgeMod, testset) # predict on test data
compare2 = cbind (actual=testset$mpg, predicted1)

mean (apply(compare2, 1, min)/apply(compare2, 1, max))

summary(linRidgeMod)

RMSE = sqrt(mean((testset$mpg-predicted1)^2))
RMSE
```

*Insight *

        - The accuracy has increased from 75% to 88%.
        - Root-Mean-Square Error has decreased from 3.242 to 1.389


