---
title: "Regression Analysis"
author: "MANOJ KUMAR - 2048015"
date: "30/01/2021"
output: word_document
---


## 1.Import the dataset data_marketing_budget_mo12 and do the exploratory data analysis .
```{r}
# Importing the shared Marketing Budget Data.

RocketData <- read.csv("data-marketing-budget.csv")
RocketData

# Summary
summary(RocketData)

```

## 2. Use Scatter Plot To Visualise The Relationship.
```{r}
scatter.smooth(x = RocketData$Month, 
               y = RocketData$Spend, 
               xlab = "Month", ylab = "Spending",  # X,Y-Lab
               main = "Spending over Month",       # Title
               col.main = "Black",                 # Title lab color
               cex.lab = "1"                       # X,Y-axis Lab color
)
```
```{r}
scatter.smooth(x = RocketData$Month, 
               y = RocketData$Sales, 
               xlab = "Month", ylab = "Sales",   # X,Y-Lab
               main = "Sales over Month",        # Title
               col.main = "Black",               # Title lab color
               cex.lab = "1"                     # X,Y-axis Lab color 
)
```


## 3.Using BoxPlot To Check For Outliers.
```{r}
boxplot(RocketData$Sales, main = "Sales" )
```
```{r}
boxplot.default(RocketData$Spend, main = "Spending" )
```

## 4. Using Density Plot To Check If Response Variable Is Close To Normal.
```{r}
library(e1071)       # For skewness function
par(mfrow = c(1,2))  # Dividing graph area in 2 columns

# Density plot for 'Spend'
plot(density(RocketData$Spend), 
     main = "Density : Spend",
     ylab = "Frequency",
     sub=paste("Skewness :",
               round(e1071::skewness(RocketData$Spend),
                     2
                     )
               )
     ) 
polygon(density(RocketData$Spend), col = "orange")

# Density plot for 'Sales'
plot(density(RocketData$Sales), 
     main = "Density : Sales",
     ylab = "Frequency",
     sub=paste("Skewness :",
               round(e1071::skewness(RocketData$Sales),
                     2
                     )
               )
     ) 

polygon(density(RocketData$Sales), col = "blue")
```


## 5. Check the Correlation Analysis.
```{r}
cor(RocketData$Spend, RocketData$Sales)
```


## 6. Build the Linear Regression Model.
```{r}
LinearRegressionModel <- lm(formula = Spend ~ Sales, data = RocketData)
LinearRegressionModel
summary(LinearRegressionModel)
```

## 7. Using p-value Check For Statistical Significance.
```{r}

```


## 8.Capture the summary of the linear model.
```{r}
ModelSummary <-summary(LinearRegressionModel)
ModelSummary
```


## 9. Also perform the Linear Diagnostics for the given data set.(Hint: plot(lmmodel) )
        
```{r}
plot(LinearRegressionModel)
```


