---
title: "Lab 11"
author: "MANOJ KUMAR"
date: "25/04/2021"
output: 
  html_document:
      toc: true # table of content true
      toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
      number_sections: false  ## if you want number sections at each table header
      theme: united  # many options for theme, this one is my favorite.
      highlight: tango  # specifies the syntax highlighting style
---

# INTRODUCTION

## 1.Load the necessary packages 

```{r}
library(tidyverse)  # For data cleaning, sorting, and visualization
library(DataExplorer) # For Exploratory Data Analysis
library(gridExtra) # To plot several plots in one figure
library(ggpubr) # To prepare publication-ready plots
library(GGally) # For correlations
library(caTools) # For classification model
library(rpart) # For classification model

```

## 2.Load the dataset

```{r}
df <- read.csv("heart.csv")
```

# EXPLORATORY DATA ANALYSIS

### 1.Tidy dataset
```{r}
copy <- df

df2 <- df %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
  ) %>%
# Recode the categorical variables as factors using the dplyr library.
  mutate(
    sex = case_when(
      sex == 0 ~ "female",
      sex == 1 ~ "male"
           ),
    fbs = case_when(
      fbs == 0 ~ "<=120",
      fbs == 1 ~ ">120"
            ),
    exang = case_when(
      exang == 0 ~ "no",
      exang == 1 ~ "yes"
            ),
    cp = case_when(
      cp == 3 ~ "typical angina",
      cp == 1 ~ "atypical angina",
      cp == 2 ~ "non-anginal",
      cp == 0 ~ "asymptomatic angina"
          ),
    restecg = case_when(
      restecg == 0 ~ "hypertrophy",
      restecg == 1 ~ "normal",
      restecg == 2 ~ "wave abnormality"
              ),
    target = case_when(
      target == 1 ~ "asymptomatic",
      target == 0 ~ "heart-disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 1 ~ "fixed defect",
      thal == 2 ~ "normal",
      thal == 3 ~ "reversable defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal)
  )

glimpse(df2) # Check that the transformnation worked
```

### 2. Missing plot
```{r}
plot_missing(df2) # Check that the transformation did not induce NA values
```
```{r}
df <- df2
```


### 3.A Visualize the data summary and distribution of each variable
```{r}
df %>%
  summary()
```

```{r}
plot_density(df, ggtheme = theme_classic2(), geom_density_args = list("fill" = "black", "alpha" = 0.6))
```

```{r}
plot_bar(df, ggtheme = theme_classic2())
```

### 4.Prepare a summary table per disease and gender

```{r}
df %>%
  group_by(target, sex) %>%
  summarise(
    n_disease = n(),
    mean_age = round(mean(age), digits=2),
    sd_age = round(sd(age), digits=2),
    mean_trestbps = round(mean(trestbps), digits=2),
    sd_trestbps = round(sd(trestbps), digits=2),
    mean_chol = round(mean(chol), digits=2),
    sd_chol = round(sd(chol), digits=2),
    mean_thalach = round(mean(thalach), digits=2),
    sd_thalach = round(sd(thalach), digits=2),
    mean_oldpeak = round(mean(oldpeak), digits=2),
    sd_oldpeak = round(sd(oldpeak), digits=2)
  )
```


# DATA VISUALIZATION

## 1.Visualization of variables per gender

## Correlation

```{r}
df <- copy %>%
  filter(
    thal != 0 & ca != 4 # remove values correspondind to NA in original dataset
    )

# ggcorr(df, palette = "RdBu")
GGally::ggcorr(df, geom = "circle")
```

```{r}
select2 <- df %>%
  dplyr::select(
    target,
    slope,
    thalach,
    restecg,
    cp
  )

ggcorr(select2, geom = "circle")
```

```{r, warning= "false"}
ggpairs(select2)
```

From the correlation study it seems that the parameters
* cp
* restecg
* thalach
* slope
are the most usefull to predict the risk of heart disease

From the EDA anlysis it seems that
* age
* sex
* cholesterol
* restecg, are also usefull

For prediction the following variables seems the most usefull
* age
* sex
* cholesterol
* restecg
* cp
* thalach
* slope

```{r}
select2
```

